---
title: "The effects of long-term trends"
author: "Robert W Schlegel"
date: "`r Sys.Date()`"
description: "This vignette investigates the effects of long-term trends on MHW detection."
# output: word_document
output: html_document
bibliography: bibliography.bib
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.align = 'center',
                      echo = TRUE, warning = FALSE, message = FALSE, 
                      eval = TRUE, tidy = FALSE)
options(scipen=999)
```

## Overview

The last major variable we want to look at that could affect MHW detection is the long-term (secular) trend present in a time series. To do so we will be using the same de-trended data from the previous two vignettes and slowly add in linear trends from 0.01 to 0.30 째C/decade in 0.0.1 째C/decade steps.

```{r r-init-quiet, echo=FALSE}
.libPaths(c("~/R-packages", .libPaths()))
library(tidyverse, lib.loc = "~/R-packages/")
library(heatwaveR, lib.loc = "~/R-packages/")
library(lubridate)
library(ncdf4)
library(doParallel)
source("code/functions.R")
```

```{r r-init, eval=FALSE}
# The packages used in this analysis
library(tidyverse)
library(heatwaveR)
library(lubridate)
library(ncdf4)
library(doParallel)

# The custom functions written for the analysis
source("code/functions.R")
```

```{r de-trend-func, eval=FALSE}
# The function used to de-trend the time series:
detrend <- function(df){
  resids <- broom::augment(lm(temp ~ t, df))
  res <- df %>%
    mutate(temp = round((temp - resids$.fitted),2))
  return(res)
}
```


## Trending

The following chunks contain the code used to systematically add increasing linear trends to the three reference time series.

```{r trending, eval=FALSE}
control_trend <- function(rate, df){
  daily_step <- rate/3652.5
  res <- df %>%
    mutate(row_index = 1:n(),
           temp = temp + (row_index*daily_step),
           test = "trend",
           index_vals = rate) %>%
    dplyr::select(test, index_vals, t, temp)
  return(res)
}
```


## Analysis

The full analysis on the results, including the functions shown above, is run for all of the tests (time series length, missing data, and long-term trend) all at once to ensure consistency of methodology across tests. For this reason the exact step-by-step process for the long-term trend analysis is not laid out below. To investigate the source code one may go to 'code/workflow.R' in the GitHub repository. A link to that site may be found in the top right of the navigation bar for this site (the GitHub icon).

```{r length-analysis-quiet, echo=FALSE}
sst_ALL_results <- readRDS("data/sst_ALL_results.Rda")
```


```{r length-analysis, eval=FALSE}
sst_ALL <- rbind(mutate(sst_WA, site = "WA"),
                 mutate(sst_NW_Atl, site = "NW_Atl"),
                 mutate(sst_Med, site = "Med"))
system.time(
  sst_ALL_results <- plyr::ddply(sst_ALL, c("site"), single_analysis, .parallel = T,
                                 full_seq = T, clim_metric = F, count_miss = T, windows = T)
) # 65 seconds
```


## Results

A more thorough explanation of the results may be found in the manuscript. Below we show what the simple results calculated above for the effect of long-term trends on the maximum intensity of the average MHWs look like.

```{r length-res}
sst_ALL_results %>% 
  filter(test == "trend",
         var == "intensity_max",
         id == "mean_perc") %>% 
  ggplot(aes(x = index_vals, y = val, colour = site)) +
  geom_line() +
  labs(x = "Long-term trend (째C/decade)", y = "Change from control (% max intensity (째C))", colour = "Site")
```


## References

